Hexagonal binning algorithm

### 2024-07-15 supplementary/hexagon_binning_algorithm.qmd
Update section lifting model to p-D
Update with quollr 1.4 version

```{r}
# umap_scurve <- read_rds(file = "data/s_curve/s_curve_umap.rds") 
# 
# scurve_scaled_obj <- gen_scaled_data(
#   data = umap_scurve)
# 
# umap_scurve_scaled <- scurve_scaled_obj$scaled_nldr
# lim1 <- scurve_scaled_obj$lim1
# lim2 <- scurve_scaled_obj$lim2
# r2 <- diff(lim2)/diff(lim1)
# 
# a1 <- calc_bins_y(bin1 = 6, q = 0.1, r2 = r2)$a1
r2 <- 1.643542
a1 <- 0.2087578
```

```{r}
all_centroids_df <- gen_centroids(bin1 = 6, r2 = r2)
```

```{r}
hex_grid <- gen_hex_coord(all_centroids_df, a1 = a1)
```

```{r}
hex_param_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354", size = 0.9) +
    geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
    geom_rect(aes(xmin = 0 - 0.03, xmax = 1 - 0.03, ymin = 0 - 0.05, ymax = r2 - 0.05), fill = "white", color = "black", alpha = 0, linewidth = 0.7) +
    geom_segment(x = 0 - 0.03, y = -0.35, xend = 1 - 0.03, yend = -0.35,
                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+
    annotate("text", x=0.5, y=-0.45, label= "r1", color = "black") +
    geom_segment(x = -0.25, y = 0 - 0.05, xend = -0.25, yend = r2 - 0.05,
                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ 
    annotate("text", x=-0.35, y=1, label= "r2", color = "black") +
    geom_segment(x = -0.1 + 0.2087578, y = -0.15, xend = -0.1 + 0.2087578*2, yend = -0.15,
                   arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a1 = 0.2087578
    annotate("text", x=0.2, y=-0.25, label= "a1", color = "black") +
    geom_segment(x = -0.15, y = -0.1*r2 + 0.1807896*2, xend = -0.15, yend = -0.1*r2 + 0.1807896*3,
                     arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a2 = 0.1807896
    annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
    annotate("text", x=-0.18, y=-0.45, label= "(s1, s2)", color = "black") ## starting point
```

```{r}
#| label: fig-hex-param
#| fig-cap: "Notations for hexagonal grid configurations."
#| fig-pos: H
#| fig-width: 2
#| fig-height: 3

@@

```{r}
all_centroids_df <- gen_centroids(bin1 = 3, r2 = r2)

hex_grid <- gen_hex_coord(all_centroids_df, a1 = 0.5218947)

>>>>

```{r}
umap_scurve <- read_rds(file = here::here("data/s_curve/s_curve_umap.rds"))

scurve_scaled_obj <- gen_scaled_data(
  data = umap_scurve)

umap_scurve_scaled <- scurve_scaled_obj$scaled_nldr
lim1 <- scurve_scaled_obj$lim1
lim2 <- scurve_scaled_obj$lim2
r2 <- diff(lim2)/diff(lim1)

sc_ltr_pos <- c(0.08, 0.96)
sc_xlims <- c(-0.5, 1.47)
sc_ylims <- c(-0.32, 2.1)
```

<!--Full hexagon grid with UMAP data-->

```{r}
## Compute hexbin parameters
num_bins_x_scurve <- 7

## hexagon binning to have regular hexagons
hb_obj_scurve <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = num_bins_x_scurve, 
  r2 = r2)

## Data set with all centroids
all_centroids_df <- hb_obj_scurve$centroids

## Generate all coordinates of hexagons
hex_grid <- hb_obj_scurve$hex_poly
```

```{r}
# Code to draw illustration for notation
start_pt <- all_centroids_df |> 
  filter(hexID == 1)
d_rect <- tibble(x1min = 0, 
                 x1max = 1,
                 x2min = 0,
                 x2max = r2)

c_start <- tibble(s1 = 0.03, s2 = -0.05)

a1 <- tibble(x = all_centroids_df$c_x[4],
             xend = all_centroids_df$c_x[5],
             y = all_centroids_df$c_y[21],
             yend = all_centroids_df$c_y[21],
             label = expression(a[1]))
a2 <- tibble(x = all_centroids_df$c_x[27],
             xend = all_centroids_df$c_x[27],
             y = all_centroids_df$c_y[27],
             yend = all_centroids_df$c_y[33],
             label = expression(a[2]))
hex_param_vis <- ggplot() + 
    geom_polygon(data = hex_grid, 
                        aes(x = x, 
                            y = y, 
                            group = hex_poly_id),
                 fill = "white", 
                 color = "#bdbdbd") +
    geom_point(data = all_centroids_df, aes(
      x = c_x, 
      y = c_y), 
      color = "#31a354", size = 0.9) +
    geom_point(data = start_pt, aes(x = c_x, 
                                    y = c_y), 
               color = "black") + 
    geom_rect(data=d_rect, 
              aes(xmin = x1min - c_start$s1,# - c_start$s1, 
                  xmax = x1max - c_start$s1,# - c_start$s1, 
                  ymin = x2min - c_start$s2,# - c_start$s2, 
                  ymax = x2max - c_start$s2),# - c_start$s2), 
              fill = "white", 
              color = "black", 
              alpha = 0, 
              linewidth = 0.7) +
    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
                                y=x2min - c_start$s2)) + 
    geom_point(data=d_rect, aes(x=x1max - c_start$s1, 
                                y=x2min - c_start$s2)) + 
    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
                                y=x2max - c_start$s2)) + 
    annotate("text", x=d_rect$x1min - c_start$s1, 
                     y=d_rect$x2min - c_start$s2,
                     label = "(0,0)", 
             hjust=-0.1, vjust=-0.3) + 
    annotate("text", x=d_rect$x1max - c_start$s1, 
                     y=d_rect$x2min - c_start$s2,
                     label = "(0,1)", 
             hjust=1.1, vjust=-0.3) + 
    annotate("text", x=d_rect$x1min - c_start$s1, 
                     y=d_rect$x2max - c_start$s2,
                     label = expression(group("(", 
                        list(0, y[2][max]),")")), 
            hjust=-0.1, vjust=1.2) + 
    geom_segment(data=d_rect, aes(
      x = x1min  - c_start$s1, # 0 - 0.03, 
      y = -0.35, 
      xend = x1max - c_start$s1, #1 - 0.03, 
      yend = -0.35), #-0.35),
      arrow = arrow(length = unit(0.03, "npc"),
                               ends = "both"), 
                 color = "black")+
    annotate("text", x=0.5, y=-0.45, 
             label = expression(r[1]), color = "black") +
    geom_segment(data=d_rect, aes(
      x = -0.25, 
      y = x2min - c_start$s2, #0 - 0.05, 
      xend = -0.25, 
      yend = x2max - c_start$s2), #r2 - 0.05),
      arrow = arrow(length = unit(0.03, "npc"),
                       ends = "both"), 
                 color = "black")+ 
    annotate("text", x=-0.35, y=1, 
             label = expression(r[2]), color = "black") +
    geom_segment(data = a1, aes(
      x = x, #-0.1 + 0.2087578, 
      y = y, #-0.15, 
      xend = xend, #-0.1 + 0.2087578*2, 
      yend = yend), #-0.15),
      arrow = arrow(length = unit(0.03, "npc"),
        ends = "both"), 
        color = "black")+ # a1 = 0.2087578
    annotate("text", 
             x=(a1$x+a1$xend)/2, 
             y=a1$y, 
             label = expression(a[1]), 
             color = "black",
             vjust = 1.2) +
    geom_segment(data = a2, aes(
      x = x, #-0.15, 
      y = y, #-0.1*r2 + 0.1807896*2, 
      xend = xend, #-0.15, 
      yend = yend), #-0.1*r2 + 0.1807896*3),
      arrow = arrow(length = unit(0.03, "npc"),
                               ends = "both"), 
      color = "black") + # a2 = 0.1807896
    annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
             label = expression(a[2]), 
             color = "black", hjust=-0.2) +
    annotate("text", x=-0.18, y=-0.25, 
      label = expression(group("(", list(s[1], s[2]), ")")),
      color = "black") 

```{r}
#| label: fig-hex-param
#| fig-cap: "The components of the hexagon grid illustrating notation."
#| out-height: 30%
#| fig-width: 2.5
#| fig-height: 3.3
#| fig-pos: H

@@

```{r}
all_centroids_df <- gen_centroids(bin1 = 4, r2 = r2, q = 0.1)

hex_grid <- gen_hex_coord(all_centroids_df, a1 = 1.043789)
